---
title: "Timestamps in dbMapR"
author: "Suman Kumar Pramanik"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Timestamps in dbMapR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
library(dbMapR)
```

## Requirement

Time stamping the time of row insertion is required to look for any merge conflict, when that row is to be updated at a later time. It becomes more relevant in a web based, multiple access and asynchronous database, where multiple users are accessing and changing a particular row in the database. R does not have any robust timestamp mechanism. Below, I give the details of using the time stamping mechanism built for this package.

## Structure of the timestamp

Timestamps are basically a `character` with following specific format, to which a `class` attribute has been added to `t_stamp`.

```
yyyymmddhhmmss{may be a decimal mark followed by upto 6 digits}@@Olson_compatible_timezone
```

Example: `20151030123634.998@@UTC`

## `cur_timestamp`

We can make current timestamp using `cur_timestamp` function. It is the only way to create a timestamp (of `t_stamp` class). This makes sure that we cannot make timestamps corresponding to any other time point other than current time point. `cur_timestamp` function uses `Sys.time` function to get the current time point. We will have to add a time zone (through `tz` parameter), which should be one of the names mantioned in `OlsonNames()` function. Do check `OlsonNames()` to see for the valid `tz` names.

## Coercion to other classes

`as.character` coerces the `t_stamp` object to `character`. The timestamp is stored in database as `character`.

`as.POSIXct` coerces the `t_stamp` object to `POSIXct` class.

`t_stamp.character` function coerces a `character` object into a `t_stamp` object given the format is conforming with `t_stamp` class as checked by the utility function `check_format_tstamp`.

## Comparision between two timestamps

`%earlier%`, `%later%` and `%same_time%` are the function which compares two `t_stamp` objects for their chronological order. They return `boolean` value.

## Examples

### Making new timestamp

```{r}
ts1 <- cur_timestamp(digits = 3L, tz = "Asia/Kolkata")
```

### Converting timestamp into POSIXct

```{r}
as.POSIXct(ts1)   # keep default tz
as.POSIXct(ts1, tz = "UTC")   # changing tz to UTC
```

### Checking for chronological order of two timestamp objects

```{r}
ts2 <- cur_timestamp(tz = "UTC")
ts3 <- cur_timestamp()

ts2 %later% ts3         # TRUE
```

`ts2` is occuring before `ts3`, but it is occuring at `UTC`, whereas `ts3` is occuring at `Asia/Kolkata` time zone (UTC - 5.5 hours). So, `ts2` is occuring later than `ts3`.

